---
title: "Rationale"
output: html_document
---

### 140201, Mass spectrometry

#### Input data
Score and coverage data associated to protein accession number come form Susann Mudrich.

* Four experimental conditions:
    + Control
    + IL27
    + Calprotectin
    + IL27 + Calprotectin
* The time series consists in  
    + 6h
    + 12h
    + 24h
* 9 biological replicates

```{r, echo=FALSE, message=FALSE}
library("ggplot2")
library("dplyr")
source("../utils.R") 
mass <- read.table("../mass.tsv", sep = "\t", header = TRUE) %>%
          tbl_df()
```

A summary of the row data is available below:

```{r, echo=FALSE}
summary(mass)
```

***

Unique accession numbers are distrubted across replicates as follows:

```{r, echo=FALSE}
# count Accession per replicates
mass2  <- group_by(mass, Accession, Condition, Time) %>%
                mutate(n = n())
# get the cumulative count of prot ID 
cumcount = c()
for(cond in levels(mass$Condition)){
    # do the cumulative sum from 9 to 1 replicate
    tab <- as.data.frame(cumsum(rev(table(distinct(filter(mass2, Condition == cond))$n)))) %>%
        add_rownames()
    colnames(tab) <- c("Replicate", "cumCount")
    cumcount <- rbind(cumcount, cbind(tab, Condition = cond))
}

# plot common proteins found in _n_ replicates
b <- ggplot(cumcount)+
    geom_bar(aes(x = Replicate, y = cumCount, fill = Condition), 
             stat = "identity", position = "dodge")+
    theme_bw()+
    ylab("Count of unique Accession")+
    xlab(expression("ID found in at least "*italic(n)* " replicate(s)"))+
    ggtitle("Time = 6h")
plot(b)
```

***

#### Filtering for protein presence across replicates 

Filtering can be dynamically be changed. By default, one protein must be found in at least `5` replicates for each condition. This means that it must be found associated with at least `20` Scores for each of the `4` conditions.  
We notice that replicate 7 for IL27 reported twice as less proteins than the other replicates.
Moreover, it appeared as an outlier and then was subsequently discarded. 

#### Imputation of missing values

Only `90` proteins are common to all replicates for all conditions. If the above filtering is less stringent, missing data are introduced. Those `NA` can be impute from neighbors data using the R bioconductor package `impute` and the function `impute.knn`.  
This function is used with default arguments displayed below:

```{r, eval=FALSE}
# Usage
impute.knn(data ,k = 10, rowmax = 0.5, colmax = 0.8, maxp = 1500, rng.seed=362436069)
# data  
An expression matrix with genes in the rows, samples in the columns
# k	
Number of neighbors to be used in the imputation (default=10)
# rowmax	
The maximum percent missing data allowed in any row (default 50%). For any rows with more than rowmax% missing are imputed using the overall mean per sample.
# colmax	
The maximum percent missing data allowed in any column (default 80%). If any column has more than colmax% missing data, the program halts and reports an error.
# maxp	
The largest block of genes imputed using the knn algorithm inside impute.knn (default 1500); larger blocks are divided by two-means clustering (recursively) prior to imputation. If maxp=p, only knn imputation is done.
```

#### Variance stabilization

It is performed using the the R package `vsn` and the function `vsn2`. The variance can be visualized before and after correction.

#### Contrast analysis

The constrat is derived from the RNA-seq analysis and uses the R bioconductor package `limma`. 
And code from Enrico Glaab:
```{r}
# code from Enrico Glaab
contrasts_limma <- function(dat, groups, ID, control, treatment){

  design <- model.matrix(~ -1 + 
                        factor(letters[ifelse(groups[c(which(groups == control),
                               which(groups == treatment))] == control, 0, 1) + 1]))
  
  outnum <- as.vector(ifelse(groups[c(which(groups == control),
                                      which(groups == treatment))] == control,
                             0, 1))
  colnames(design) <- unique(letters[outnum + 1])
  
  corfit <- duplicateCorrelation(dat[,c(which(groups == control),
                                        which(groups == treatment))],
                                 design,
                                 block = ID)
  #corfit$consensus
  fit <- lmFit(dat[,c(which(groups == control),
                         which(groups == treatment))],
               design,
               block = ID,
               correlation = corfit$consensus)
  
  cm <- makeContrasts(comp = b - a, levels = design)
  fit2 <- contrasts.fit(fit, cm)
  return(eBayes(fit2))
}
```

#### Output tables

Be default, the top 20 entries, sorted by the adjusted p-values are displayed.
